---
title: "3_Fishing_inspection"
author: "Adel"
date: "09/08/2021"
output: html_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
### libraries
# Load fishwatchr
library(fishwatchr)
library(tidyverse)
library(glue)
library(rnaturalearthhires)
library(rgeos)
### 

###
# Establish connection to BigQuery project
con <- DBI::dbConnect(drv = bigrquery::bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)
###

```

## Question 1
1.	We may also want to know about fishing. Plot a map of the track and fishing points by the vessel with the MMSI 367650000 between March 1 2017 and March 5 2017. Are there any issues with the track?

```{r fishing track query}

##Q Notes -- I took this query from the fishwatchr documentation, updated (DATE to timestamp) and adapted (changed timestamp to _partitiontime) and this reduced the cost of the query from ~ 782 GB to 7 GB
### query

track <- c("

SELECT 
  ssvid
	,lon
	,lat
	,TIMESTAMP
	,speed_knots
	,- 1 * elevation_m AS depth_m
	,distance_from_shore_m / 1000 AS distance_from_shore_km
	,nnet_score

FROM `gfw_research.pipe_v20201001_fishing` 
WHERE seg_id IN (
		SELECT seg_id
		FROM `gfw_research.pipe_v20201001_segs`
		WHERE overlapping_and_short IS FALSE
			AND good_seg IS TRUE
			)
			
	AND ssvid = '367650000'
	AND _partitiontime >= TIMESTAMP('2017-03-01') AND _partitiontime <= TIMESTAMP('2017-03-05')
	AND nnet_score IS NOT NULL
ORDER BY TIMESTAMP

")

```

```{r run query}

## run query
clean_track<-gfw_query(query = track, 
                          run_query = TRUE, 
                          con = con)
```

```{r plot track}
## inspect track


clean_track$data$ssvid<-as.factor(clean_track$data$ssvid)
clean_track$data$timestamp<-clean_track$data$TIMESTAMP ### need to do this for plot function to run

fishing_track_review(track_df = clean_track$data,
                     color_fishing = TRUE,
                     globe_location = 'lowerright',
                     time_series_labels = c('speed_knots','depth_m','distance_from_shore_km'),
                     vessel_name = 'Alaska Endeavor')

```
This is a small-ish trawler (37m) -- and it's estimates of fishing segments seems really high. Are they inverted with the steaming segs? 

## Question 2
2.	How many hours of fishing are estimated during this time? Is it different if you use the `pipe_production_v20201001.proto_events_fishing` table?

```{r fishing hours from research pipe}
##
hoursq<- c("
---------------------------------------
--  Note: I took the query from example queries https://github.com/GlobalFishingWatch/bigquery-documentation-wf827/blob/f546b4efd638230eb83ad61013ffb4ad7e74f290/queries/examples/2019/Q4/fishing_effort/research-table-examples-1.sql
-- and updated to have the partition time

-- How many hours did ssvid 123456789 fish on January 1, 2019,
-- in segments that are likely not noise? (use `good_segs`)
-------------------------------------

WITH
  good_segments AS (
  SELECT * FROM
    `gfw_research.pipe_v20201001_segs`
  WHERE good_seg)
--
--
SELECT
  sum(fishing_hours) fishing_hours
FROM
  `gfw_research.pipe_v20201001_segs_daily`
WHERE
  ssvid = '367650000'
  and _partitiontime >= timestamp('2017-03-01') AND _partitiontime <= timestamp('2017-03-05')
  and seg_id in (select seg_id from good_segments)

")

hours<-gfw_query(query = hoursq, 
                          run_query = TRUE, 
                          con = con)

```

98 fishing hours from estimate that uses good segs and research pipe

```{r fishing hours from pipe_production_v20201001.proto_events_fishing}
## I couldn't find an example query for fishing events in git hub examples, slack or the slides

```

## Question 3
3.	What type of fishing vessel is it? 

```{r vessel info}
#

tq<- c(
"
SELECT 
ssvid,
inferred.inferred_vessel_class_ag,
registry_info.best_known_vessel_class,
best.best_vessel_class,
on_fishing_list_known,
on_fishing_list_nn,
on_fishing_list_sr,
on_fishing_list_best
FROM `world-fishing-827.gfw_research.vi_ssvid_v20210706`
WHERE ssvid = '367650000'

"
)


## run query
dd<-gfw_query(query = tq, 
                          run_query = TRUE, 
                          con = con)

view(dd$data)

```

This is a trawler

## Question 4
4.	Provide the other vessel identity information and registry records during this time.

Note this query returns no records...
```{r vessel database information}

tq<- c(
"
---SET your date minimum of interest
CREATE TEMP FUNCTION minimum() AS (DATE("2017-03-01"));
---SET your date maximum of interest
CREATE TEMP FUNCTION maximum() AS (DATE("2017-03-05"));
--SET your year of interest
CREATE TEMP FUNCTION yoi() AS (CAST(2018 AS INT64));

--####
SELECT
auth_ssvid,
auth_shipname,
auth_imo,
auth_flag,
registry_authorized_to,
registry_authorized_from,
reg,
first_timestamp,
last_timestamp
FROM(
 SELECT
  identity.ssvid as auth_ssvid,
  shipname as auth_shipname,
  identity.imo as auth_imo,
  flag as auth_flag,
   authorized_to as registry_authorized_to,
  authorized_from as registry_authorized_from,
  is_active as registry_active,
  SAFE_CAST( SPLIT(list_uvi, '-')[OFFSET(0)] AS string) AS reg,
    scraped,
    list_uvi,
   geartype_original as registry_gear,
   feature_gear,
   last_modified,
   first_timestamp,
   last_timestamp
FROM
  `vessel_database.all_vessels_v20210601`
  LEFT JOIN UNNEST(registry)
  LEFT JOIN UNNEST(activity)
  LEFT JOIN UNNEST(feature.geartype) as feature_gear
WHERE
identity.ssvid = '367650000'
AND
authorized_to>first_timestamp
AND
authorized_from<last_timestamp
AND
DATE(last_timestamp) >= minimum()
AND
DATE(first_timestamp) <= maximum()
)
WHERE
(EXTRACT (YEAR from registry_authorized_to) = yoi()
OR
EXTRACT (YEAR from registry_authorized_from) = yoi())
GROUP BY
auth_ssvid,
auth_shipname,
auth_imo,
auth_flag,
registry_authorized_to,
registry_authorized_from,
reg,
first_timestamp,
last_timestamp
")


## run query
dd<-gfw_query(query = tq, 
                          run_query = TRUE, 
                          con = con)

view(dd$data)

```
When I run the query that doesn't specify the time frame, it looks like it was dropped from the US registry in 2016, in 2017 it wasn't a registered fishing vessel
```{r}
tq<- c(
"
SELECT
 matched, loose_match,
 identity,
 feature,
 activity,
 ARRAY(
   SELECT list_uvi
   FROM UNNEST(registry)
   GROUP BY 1 ) AS list_uvi
FROM `world-fishing-827.vessel_database.all_vessels_v20210601`
WHERE identity.ssvid = '367650000'
"
)


## run query
dd<-gfw_query(query = tq, 
                          run_query = TRUE, 
                          con = con)

view(dd$data)

```

## Question 5
5.	Let’s take the scale up for a second. Plot a raster of all fishing by Chinese flagged squid jiggers vessels in NPFC in 2018. What are some caveats to take into consideration with this data?

## Ran out of time...
