---
title: "Training Questions"
author: "Rollan Geronimo"
date: "12/14/2021"
output: 
  bookdown::html_document2: 
    toc: true
    code_folding: hide
    toc_depth: 2.0
---

```{r setup, include=FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r load libraries, include=FALSE}
library(bigrquery)
library(fishwatchr)
library(dplyr)
library(ggplot2)
library(glue)
library(tidyverse)
library(raster)
library(sf)
library(gt)

BQ_connection <-  dbConnect(bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)
con <- DBI::dbConnect(drv = bigrquery::bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)
```

# Fishing Inspection Questions

## Fishing track and points plot
**1.	We may also want to know about fishing. Plot a map of the track and fishing points by the vessel with the MMSI 367650000 between March 1 2017 and March 5 2017. Are there any issues with the track?**

There are tracks labelled as apparent fishing but looks like non-fishing (e.g., cruising / transiting) patterns. 

```{r fishingInspection_Q1, echo = TRUE, message = FALSE, warning = FALSE}

ssvid = '367650000'
dateRange <- c("2017-03-01","2017-03-05")

# ssvid = '417000121'
# dateRange <- c("2020-01-01", "2020-12-31")
query_Tracks <- glue::glue(
  " SELECT
    ssvid,
    seg_id,
    lon,
    lat,
    hours,
    timestamp,
    nnet_score
    FROM
    `gfw_research.pipe_v20201001_fishing` 
    WHERE seg_id IN (
      SELECT seg_id 
      FROM `gfw_research.pipe_v20201001_segs` 
      WHERE 
      overlapping_and_short IS FALSE AND 
      good_seg IS TRUE) AND
    ssvid = '{ssvid}' AND
    _partitiontime >= TIMESTAMP('{dateRange[1]}') AND 
    _partitiontime <= TIMESTAMP('{dateRange[2]}') AND
    # Should we include nnet_scores that are NULL?
    nnet_score IS NOT NULL 
    ORDER BY timestamp"
)

vTracks <- gfw_query(query = query_Tracks, 
                          run_query = TRUE, 
                          con = con)$data

plot_vessel_track_fish(
  vTracks,
  bounding_box_man = NA,
  color_fishing = TRUE,
  color_by = "nnet_score",
  map_res = 10,
  theme = "light",
  buffer = 2,
  add_gridlines = TRUE,
  gfw_proj_name = "Equal Earth",
  vessel_name = "Alaska Endeavor" # name obtained by searching for the MMSI in the GFW map
)

# # This is another way to plot the tracks
# land_sf <- rnaturalearth::ne_countries(scale = 10, returnclass = 'sf')
# plot_vessel_track(
#   vTracks,
#   lat = "lat",
#   lon = "lon",
#   show_fishing = TRUE,
#   spatial = TRUE
# )

```


**2.	How many hours of fishing are estimated during this time? Is it different if you use the `pipe_production_v20201001.proto_events_fishing` table?**

The proto_events_fishing table will give a slightly restrictive fishing (based on consistency and duration filter).

```{r fishingInspection_Q2, echo = TRUE, message = FALSE, warning = FALSE}
q_events <- glue::glue(
  " SELECT
    SUM(IF(event_type = 'fishing', 
     TIMESTAMP_DIFF(event_end, event_start, MINUTE),
     0))/60 AS fishing_hours
    FROM
    `pipe_production_v20201001.proto_events_fishing` 
    WHERE
    event_start BETWEEN TIMESTAMP('{dateRange[1]}') AND TIMESTAMP('{dateRange[2]}') 
    AND
    seg_id LIKE '{ssvid}-%' # But seg_id does not always rely only on ssvid. It can sometimes be imo. 
    "
) # better to get the seg_id and match with the seg_id in the proto table

# "Fishing hours" not exactly interpreted as gear-in-the-water time but more broadly in terms of vessel behavior that appears as fishing operations.
# GFW map now uses proto_fishing_events instead of pipe_fishing 
# Can use proto_fishing_events for single vessel analysis
# Use fishing days but put note on what is counted as 'fishing day'
# JSON_EXTRACT_SCALAR(event_vessels,"$[0].ssvid") as ssvid

events_hours <- gfw_query(query = q_events, 
                          run_query = TRUE, 
                          con = con)$data
```

The total fishing hours is `r round(sum(vTracks[vTracks$nnet_score == 1, "hours"]))` out of `r round(sum(vTracks$hours))` total voyage hours using the *'gfw_research.pipe_v20201001_fishing'* table. 

The total fishing hours using the *'pipe_production_v20201001.proto_events_fishing'* is round(`r events_hours`).


**3.	What type of fishing vessel is it?**

 

```{r fishingInspection_Q3, echo = TRUE, message = FALSE, warning = FALSE}
q_vesselClass <- glue::glue(
  "SELECT 
    year, 
    inferred.inferred_vessel_class AS inferred_class, 
    registry_info.best_known_vessel_class AS registry_class,
    best.best_vessel_class AS best_class
  FROM `world-fishing-827.gfw_research.vi_ssvid_byyear_v20210706` 
  WHERE ssvid = '{ssvid}' AND year = 2017
   "
)

v_class <- gfw_query(query = q_vesselClass, 
                          run_query = TRUE, 
                          con = con)$data

```
Best vessel class = `r v_class$best_class`.

**4.	Provide the other vessel identity information and registry records during this time.**

The vessel identity information can be derived from the vessel_database.all_vessels_v20210701. Registry information is in the vessel info table vi_ssvid_byyear (since we need to get the registry info for the given period). 

What is the difference between "normalized" and non-normalized values in the *vi_ssvid_byyear* table?

```{r fishingInspection_Q4, echo = TRUE, message = FALSE, warning = FALSE}

q_vesselInfo <- glue::glue(
  "SELECT 
     ssvid,
     year,
     registry_info.best_known_shipname,
     ais_identity.callsign,
     ais_identity.imo,
     ais_identity.shiptype,
     ais_identity.flag_mmsi,
     inferred.inferred_vessel_class,
     inferred.inferred_vessel_class_score_byyear,
     registry_info.registries_listed,
     registry_info.best_known_flag,
     registry_info.best_known_callsign,
     registry_info.best_known_imo,
     registry_info.best_known_vessel_class,
     registry_info.best_known_length_m,
     registry_info.best_known_tonnage_gt,
     registry_info.best_known_engine_power_kw,
     registry_info.best_known_crew_size,
     best.registry_net_disagreement
  FROM `world-fishing-827.gfw_research.vi_ssvid_byyear_v20210913` 
  WHERE ssvid = '{ssvid}' AND year = 2017 
   "
)
 
v_info <- gfw_query(query = q_vesselInfo, 
                          run_query = TRUE, 
                          con = con)$data

v_info %>% t() %>% data.frame() %>% rownames_to_column(var="parameters") %>% rename(values = '.') %>% 
  gt() %>%
  tab_style(
    style = cell_text(color = "black", weight = "bold"),
    locations = list(
      cells_row_groups(),
      cells_column_labels(everything())
    )
  ) %>% 
  cols_label(
    parameters = "Parameters",
    values = "Values"
  )

```

**5.	Let’s take the scale up for a second. Plot a raster of all fishing by Chinese flagged squid jiggers vessels in NPFC in 2018. What are some caveats to take into consideration with this data?**

I used the 'proto_events_fishing_v20211212' table. 

```{r fishingInspection_Q5, echo = TRUE, message = FALSE, warning = FALSE}

q_effort <- glue::glue(
  "WITH 
positions AS (
          SELECT
          ssvid,
          lat,
          lon,
          nnet_score,
          night_loitering,
          hours
          FROM `world-fishing-827.gfw_research.pipe_v20201001_fishing`,
          UNNEST(regions.rfmo) rfmo
          WHERE EXTRACT(DATE FROM _PARTITIONTIME) BETWEEN ('2018-01-01') AND ('2018-12-31')
              AND seg_id IN (
                  SELECT seg_id
                  FROM `world-fishing-827.gfw_research.pipe_v20201001_segs`
                  WHERE good_seg
                  AND NOT overlapping_and_short
              )
              AND rfmo = 'NPFC'
      ),

fishing_vessels AS (
          SELECT
              ssvid,
              year,
              best_vessel_class AS geartype,
              best_flag as flag
          FROM `world-fishing-827.gfw_research.fishing_vessels_ssvid_v20210913` 
          WHERE year = 2018
          AND best_flag = 'CHN'
          AND best_vessel_class = 'squid_jigger'
      ),

positions_vi AS (
    SELECT * EXCEPT(nnet_score, night_loitering),
        IF(b.geartype = 'squid_jigger', night_loitering, nnet_score) as nnet_score
    FROM positions a
    LEFT JOIN fishing_vessels b
    USING (ssvid)
),

grouped AS (
    SELECT 
        FLOOR(lat * 10)/10 as lat_bin,
        FLOOR(lon * 10)/10 as lon_bin,
        SUM(IF(nnet_score > 0.5, hours, 0) ) AS fishing_hours
    FROM positions_vi
    GROUP BY 1,2
)

SELECT * FROM grouped
   "
)

fishing_effort <- gfw_query(query = q_effort, 
                          run_query = TRUE, 
                          con = con)$data

fishing_effort %>% filter(fishing_hours>1) %>%
    ggplot() +
    geom_raster(aes(x = lon_bin,
                    y = lat_bin,
                    fill = fishing_hours)) +
    geom_gfw_eez(center = 160) +
    geom_gfw_land(center = 160) +
    #coord_sf(crs = CRS('+proj=longlat +datum=WGS84'), expand=F)+
#    lims(x = lonRange, y = latRange) +
    scale_fill_gradientn(colors = gfw_palette('map_effort_dark'),
                         limits = c(0,500),
                         oob = scales::squish,
                         na.value = NA) +
    labs(fill = 'Fishing hours') +
    theme_gfw_map() +
    theme(axis.text = element_text(family = "Roboto", color = "#848b9b", 
              size = 10),
          legend.text = element_text(family = "Roboto", 
              color = "#848b9b", size = 10), legend.title = element_text(family = "Roboto Bold", 
              color = "#363c4c", size = 10), plot.title = element_text(family = "Roboto Bold", 
              color = "#363c4c", size = 12), plot.subtitle = element_text(family = "Roboto", 
              color = "#363c4c", size = 12)
          )

```
