---
title: "Fishing inspection"
author: "Edaysi Bucio"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Establish connection to BigQuery project

```{r}
con <- DBI::dbConnect(bigrquery::bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)

#Run this line if you receive an error about your BigQuery credentials
#bq_auth() 

```



## Download packages
```{r}
devtools::install_github("ropensci/rnaturalearthhires") #If problems isntalling the package, look at this https://github.com/ropensci/rnaturalearthhires/issues/1

#Install fishwatchr
devtools::install_github("GlobalFishingWatch/fishwatchr")
library(fishwatchr)
library(tidyverse)
library(bigrquery)
library(ggsci) 
library(rnaturalearth)
library(rnaturalearthhires)
library(sf)
library(raster)
library(viridis)
library(googledrive)
library(sp)
library(devtools)

```


Fishing inspection

**1. We may also want to know about fishing. Plot a map of the track and fishing points by the vessel with the MMSI 367650000 between March 1 2017 and March 5 2017. Are there any issues with the track?**

```{sql connection = con, output.var = "track_inspection"}

WITH
vessel AS (
SELECT
  *
FROM (
  SELECT
    timestamp,
    lat,
    lon,
    ssvid,
    seg_id,
    nnet_score,
    hours,
  IF(nnet_score > 0.5, hours,0) fishing_hours
  FROM
    `world-fishing-827.gfw_research.pipe_v20201001`
  WHERE
    # Select data range for track using date to make query cheaper
    _partitiontime >= TIMESTAMP("2017-03-01") 
    AND  _partitiontime <= TIMESTAMP("2017-03-05")
    # List of MMSI(s) to get tracks for
    AND ssvid IN ("367650000") )
# only select points that are part of non-noise segments (good_seg)
WHERE
  seg_id IN (
  SELECT
    seg_id
  FROM
    `gfw_research.pipe_v20201001_segs`
    WHERE 
   good_seg = true AND
        NOT overlapping_and_short)
  ORDER BY
  ssvid,
  timestamp   
),
identity AS (
SELECT
ssvid,
activity,
activity.first_timestamp, 
activity.last_timestamp,
ais_identity.n_shipname,
ais_identity.n_callsign,
ais_identity.n_imo
FROM
   `world-fishing-827.gfw_research.vi_ssvid_v20210706`
)
-- Return final data after joining with vessel info
SELECT 
*
FROM 
vessel
LEFT JOIN 
identity
USING (ssvid)
 ORDER BY
timestamp DESC
  

```
**1.1 Are there any issues with the track?**
Answer: I don't see any issues with the track. Looks like there is consistent fishing activity. 
```{r}

fishing_track_review(track_df = track_inspection,
                     color_fishing = TRUE,
                     globe_location = 'upperright',
                     theme = 'dark',
                     time_series_labels = NA,
                     vessel_name = 'ALASKA_ENDEAVOR')
```

**2. How many hours of fishing are estimated during this time? Is it different if you use the `pipe_production_v20201001.proto_events_fishing` table?**
Answer: With gfw_research.pipe_v20201001 there are 98.87. With pipe_production_v20201001.proto_events_fishing` there are X



**3. What type of fishing vessel is it?**
Answer: Its a trawler
```{sql connection = con, output.var = "vessel_type"}

SELECT
ssvid, 
ais_identity.n_shipname,
ais_identity.n_shipname_mostcommon,
ais_identity.n_callsign,
ais_identity.n_callsign_mostcommon,
ais_identity.n_imo,
ais_identity.n_imo_mostcommon,
ais_identity.shiptype,
inferred.inferred_vessel_class
FROM `world-fishing-827.gfw_research.vi_ssvid_v20210706` 
WHERE ssvid in  ('367650000') 

```
```{r}
View(vessel_type)
```


**4. Provide the other vessel identity information and registry records during this time.**
Answer: 

- The vessel was broadcasting with the call sign and IMO numbers WDF7174 and 793355550, then with 7DF7174 and 792797470, respectively.

- During 2017, the vessel showed registry records in the NEPAC, REV, and USA. Question: Not sure what REV or USA stands for registry records. 

```{sql connection = con, output.var = "registry_records"}
SELECT
*
FROM gfw_research.vi_ssvid_byyear_v20210706
WHERE ssvid IN ('367650000')
AND year = 2017

```


**5.Let’s take the scale up for a second. Plot a raster of all fishing by Chinese flagged squid jiggers vessels in NPFC in 2018. What are some caveats to take into consideration with this data?**

````{sql connection = con, output.var = "Chinese_squid"}
WITH
-- area of interest NPFC :: polygon to select squid chinese fishing vessels during 2018
aoi as(
  SELECT
  st_GeogFromText(string_field_1, make_valid=>TRUE) AS polygon
  FROM
  `world-fishing-827.ocean_shapefiles_all_purpose.NPFC_shape`
),



    -- Caclulate daily fishing and non-fishing hours for each vessel in the NPFC AKA Aoi. Note: Probably using the night_loitering for getting squid fishing activity
  positions AS ( 
  SELECT   
  EXTRACT(year from timestamp) AS year,
    ssvid,
    lat as latitude,
    lon as longitude,
    hours,
    nnet_score,   
    FROM
    `gfw_research.pipe_v20201001_fishing`
  WHERE
    _partitiontime BETWEEN TIMESTAMP('2018-01-01') 
    AND TIMESTAMP('2018-12-31')
),
      
 ssvid_in_aoi AS ( 
  SELECT
  * 
  FROM
    positions
  WHERE
  IF
    (ST_CONTAINS( (
        SELECT
          polygon
        FROM
          aoi),
        ST_GEOGPOINT(longitude,
          latitude)),
      TRUE,
      FALSE)
),
-- Bin data and caluclate fishing hours
fishing_gridded AS (  
SELECT
  *,
  -- Calcultate 10th degree grid cell for each position 
  floor(latitude * 10) as lat_bin,
  floor(longitude * 10) as lon_bin,
  IF(nnet_score > 0.5, hours, 0) as fishing_hours
  FROM
  ssvid_in_aoi
),
-- Summarize fishing hours by 10th degree lat/lon grid cell
total_fishing AS (
SELECT
  ssvid,
  year,
   -- Divide by 10 to convert back to decimal degrees
  lat_bin / 10 as lat_bin,
  lon_bin / 10 as lon_bin,
  SUM(hours) as hours,
  SUM(fishing_hours) as fishing_hours,
  COUNT(*) as positions
FROM fishing_gridded
GROUP BY ssvid, year, lat_bin, lon_bin
),
-- Get vessel info. Note: Here I'm trying to select just infro for vessels flagged to China and squid jiggers
vessel_info AS (
SELECT 
identity.ssvid AS ssvid,
identity.n_shipname,
feature.geartype AS squid,
a.first_timestamp,
a.last_timestamp,
identity.flag AS flag,
FROM `world-fishing-827.vessel_database.all_vessels_v20200101`, UNNEST (activity) AS a
WHERE ( matched OR loose_match )
  /* Check whether `geartype` of the vessel is
  either purse_seiner or tuan_purse_seiner */
  AND EXISTS (
    SELECT *
    FROM UNNEST (feature.geartype) AS x
    WHERE x IN ("squid_jigger") )
  AND "other_fishing" NOT IN UNNEST (feature.geartype)
  AND identity.flag = 'CHN'
)
-- Return final data after joining with vessel info
SELECT 
    ssvid,
    lat_bin,
    lon_bin,
    hours,
   fishing_hours,
   positions,
   n_shipname,
   flag,    
FROM 
total_fishing
LEFT JOIN 
vessel_info
USING (ssvid)

```

