---
title: "Training Questions Revamped by Chih-Yu"
author: "Chih-Yu Lai"
date: "11/24/2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

##Load libraries
```{r loadPackages, include = FALSE}
library(fishwatchr)
library(dplyr)
library(sf)
library(kableExtra)
library(ggplot2)
library(glue)
library(tidyverse)
library(raster)
library(DT)
library(here)
library(measurements) 
library(stringr) 
```

##Connect to BQ
```{r}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), project = "world-fishing-827", use_legacy_sql = FALSE)
```

##Set parameters
```{r setParameters}
ssvid = "367650000"
start_date = "2017-03-01"
end_date = "2017-03-05" 
```


#Fishing inspection

##1.We may also want to know about fishing. Plot a map of the track and fishing points by the vessel with the MMSI 367650000 between March 1 2017 and March 5 2017. Are there any issues with the track?

```{r, echo=FALSE}
VesselTracks <- 
  'WITH
      ########################################
      # This subquery identifies good segments
      good_segments AS (
      SELECT
        seg_id
      FROM
        `gfw_research.pipe_v20201001_segs`
      WHERE
        good_seg
        AND positions > 10
        AND NOT overlapping_and_short
      )
      ########################################
  # This subquery fishing query gets all fishing.
    SELECT
        ssvid,
        seg_id,
        lon,
        lat,
        hours,
        timestamp,
        nnet_score
    FROM
    `gfw_research.pipe_v20201001_fishing` 
    WHERE 
    _partitiontime BETWEEN TIMESTAMP("2017-03-01") and TIMESTAMP("2017-03-05")
    AND ssvid = "367650000"
    AND seg_id IN (
          SELECT
            seg_id
          FROM
            good_segments) 
    AND nnet_score IS NOT NULL 
    ORDER BY timestamp DESC
    '
VesselTracks_df<- DBI::dbGetQuery(con, VesselTracks)

fishing_track_review(track_df = VesselTracks_df,
                     color_fishing = TRUE,
                     globe_location = 'upperright',
                     theme = 'light',
                     time_series_labels = NA,
                     vessel_name = 'ALASKA_ENDEAVOR')
```

```{r}
table_df <- VesselTracks_df %>% dplyr::select(ssvid,seg_id,lon,lat,timestamp,hours,nnet_score) 
table_df %>%  arrange(desc(timestamp))%>% datatable(colnames = c("ssvid","seg_id","lat","lon","timestamp","hours","nnet_score"),caption = sprintf("Track table"),options = list(order = list(7, 'desc')))
```


##2.How many hours of fishing are estimated during this time? Is it different if you use the `pipe_production_v20201001.proto_events_fishing` table?
A:109.715833333
```{r}
FishingHour<-sum(table_df$hours)
```

```{r, echo=FALSE}
VesselTracks2 <- 
  'WITH
      ########################################
      # This subquery identifies good segments
      good_segments AS (
      SELECT
        seg_id
      FROM
        `gfw_research.pipe_v20201001_segs`
      WHERE
        good_seg
        AND positions > 10
        AND NOT overlapping_and_short
      )
      ########################################
  # This subquery fishing query gets all fishing.
    SELECT
        ssvid,
        seg_id,
        lon,
        lat,
        hours,
        timestamp,
        nnet_score
    FROM
    `pipe_production_v20201001.proto_events_fishing` 
    WHERE 
    _partitiontime BETWEEN TIMESTAMP("2017-03-01") and TIMESTAMP("2017-03-05")
    AND ssvid = "367650000"
    AND seg_id IN (
          SELECT
            seg_id
          FROM
            good_segments) 
    AND nnet_score IS NOT NULL 
    ORDER BY timestamp DESC
    '
VesselTracks2_df<- DBI::dbGetQuery(con, VesselTracks2)
table2_df <- VesselTracks2_df %>% dplyr::select(ssvid,seg_id,lon,lat,timestamp,hours,nnet_score) 
table2_df %>%  arrange(desc(timestamp))%>% datatable(colnames = c("ssvid","seg_id","lat","lon","timestamp","hours","nnet_score"),caption = sprintf("Track2 table"),options = list(order = list(7, 'desc')))
FishingHour2<-sum(table2_df$hours)
```

##3.What type of fishing vessel is it? 
A: trawlers
```{r, echo=FALSE}
VesselType <- 
  '
  SELECT
      ssvid, 
      ais_identity.n_shipname as n_shipname,
      ais_identity.n_shipname_mostcommon as shipname,
      ais_identity.n_callsign as n_callsign,
      ais_identity.n_callsign_mostcommon as callsign,
      ais_identity.n_imo as n_imo,
      ais_identity.n_imo_mostcommon as imo,
      ais_identity.shiptype as shiptype,
      inferred.inferred_vessel_class as vessel_class
  FROM `gfw_research.vi_ssvid_byyear_v20210913` 
  WHERE year = 2017
  AND   ssvid = "367650000"
  '
VesselType_df<- DBI::dbGetQuery(con, VesselType)
View(VesselType_df)
```

##4.Provide the other vessel identity information and registry records during this time.
A: This vessel had registry records in the NEPAC, REV, and USA in 2017.
```{r, echo=FALSE}
VesselIdentity <- 
  '
  SELECT
         *
  FROM `gfw_research.vi_ssvid_byyear_v20210913` 
  WHERE year = 2017
  AND   ssvid = "367650000"
  '
VesselIdentity_df<- DBI::dbGetQuery(con, VesselIdentity)
```

##5.Let’s take the scale up for a second. Plot a raster of all fishing by Chinese flagged squid jiggers vessels in NPFC in 2018. What are some caveats to take into consideration with this data?

```{r, echo=FALSE}
CHNsj2018 <- 
  'WITH
      ########################################
      # This subquery identifies good segments
      good_segments AS (
      SELECT
        seg_id
      FROM
        `gfw_research.pipe_v20201001_segs`
      WHERE
        good_seg
        AND positions > 10
        AND NOT overlapping_and_short
      ),
      
      ####################################################################
      # Get the list of active fishing vessels that pass the noise filters
      fishing_vessels AS (
      SELECT
        ssvid,
        year,
        best_flag,
        best_vessel_class
      FROM 
        `gfw_research.fishing_vessels_ssvid_v20210706`
      WHERE 
        best_flag = "CHN" 
        AND best_vessel_class = "squid_jigger"
        AND year = 2018
      ),
      
      ########################################################################
      # Read in shapefile of npfc
      npfc AS (
      SELECT
            ST_GEOGFROMTEXT(string_field_1) AS polygon
      FROM
            `world-fishing-827.ocean_shapefiles_all_purpose.NPFC_shape`
      ),

      ########################################
      # This subquery fishing query gets all fishing.
      fishing as (
      SELECT
          ssvid,
          seg_id,
          lat,
          lon,
          hours,
          EXTRACT(year FROM timestamp) as year,
          timestamp,
          nnet_score,
          night_loitering
      FROM
      `gfw_research.pipe_v20201001_fishing` 
      WHERE 
      _partitiontime BETWEEN TIMESTAMP("2018-01-01") and TIMESTAMP("2018-12-31")
      AND seg_id IN (
            SELECT
              seg_id
            FROM
              good_segments) 
      AND nnet_score IS NOT NULL 
      ORDER BY timestamp DESC
      ),
      
      ###############################################################################
      # Filter fishing to the list of active fishing vessels in that year within nfpc
      fishing_filtered AS (
              SELECT *
              FROM fishing
              JOIN fishing_vessels
              USING(ssvid, year)
      ),
      fishing_filtered_in_NFPC AS (
      SELECT
        *
      FROM
        fishing_filtered
      WHERE
      IF
        (ST_CONTAINS( (
            SELECT
              polygon
            FROM
              npfc),
            ST_GEOGPOINT(lon, lat)),
          TRUE,
          FALSE)
      ),
      ###################################################################################################
      # Create fishing_hours attribute by using night_loitering as indicator of fishing for squid jiggers
      fishing_hours_filtered AS (
      SELECT *,
          CASE
            WHEN best_vessel_class = "squid_jigger" and night_loitering = 1 THEN hours
            WHEN best_vessel_class != "squid_jigger" and nnet_score > 0.5 THEN hours
            ELSE 0
          END
          AS fishing_hours
      FROM fishing_filtered_in_NFPC
      ),
      ###################################################################################
      # This subquery sums fishing hours and converts coordinates back to decimal degrees
      fishing_binned AS (
      SELECT
        year,
        FLOOR(lat * 10) / 10 as lat_bin,
        FLOOR(lon * 10) / 10 as lon_bin,
        SUM(hours) as hours,
        SUM(fishing_hours) as fishing_hours
      FROM fishing_hours_filtered
      GROUP BY year, lat_bin, lon_bin
      )

      SELECT *
      FROM fishing_binned
    '
CHNsjs2018_df<- DBI::dbGetQuery(con, CHNsj2018)



```



```{r}
  CSJeffort_df <- CHNsjs2018_df %>% 
  filter(fishing_hours>0) %>%
  mutate(lon_bin = ifelse(lon_bin < 0, lon_bin + 360, lon_bin))

  CSJeffort_df%>% 
  recenter_raster(raster_df = .,
                  res = 0.1,
                  x_lab = "lon_bin",
                  y_lab = "lat_bin",
                  fill_lab = "fishing_hours",
                  center = 180,
                  proj = "+proj=eqearth +lon_0=0 +wktext") %>%

  ggplot() +
  geom_raster(aes(x = lon_bin,
                  y = lat_bin,
                  fill = fishing_hours)) +
  geom_gfw_land(theme = "light") +
  labs(title = "Fishing activity by Chinese flagged squid jiggers",subtitle = "2018") +
  scale_fill_gradientn(colours = gfw_palette("map_effort_dark"),
                       limits = c(0,100),
                       oob = scales::squish,
                       na.value = NA,
                       name = "Fishing Hours") +
  labs(fill = "fishing_hours")
  theme_gfw_map()
```

